Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-02-21 article/article.Rmd
+ 4 species for debugging + fixes to article so it compiles + fixes to nmds analyses so they use sparse distances

# load .rda file
session::restore.session('results/results.rda')

@@

To understand the relationship between the surrogate-based targets used to generate a prioritizations and the amount of genetic variation secured, we generated a collection of single-species prioritizations. For each species, we generated two sets of prioritizations using increasing environmental and geographic-based targets (respectively). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %. None of these prioritizations contained any amount-based targets.

@@

For each scenario, we generated three types of prioritizations. We generated prioritizations using only amount-based, both amount- and surrogate-based targets, and both amount- and genetic-based targets. The purely amount-based prioritizations represented those typically generated in conservation planning exercises and serve as our baseline. We used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX]. We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % to ensure that most of the repsective variation was secured for each species. To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$amount.replicates)` replicate prioritizations. Since all other contexts were likely to be associated with a single optimal solution, we used only a single replicate.

@@

## Results
<!-- Surrogacy analysis -->
Prioritizations generated using greater environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1). The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (`r xxx`). Generally, prioritizations needed to capture at least `r xxx` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation. Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (`r xxxx`). Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation. These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation.

<!-- Scenario analysis -->
The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2). No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$), or under any of the different scenarios (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$). The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation. Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r ` \pm `r `) and neutral (mean `r ` \pm `r `) genetic variation. These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r `$, $P `r `$) or genetic-based targets ($Z = `r `$, $P `r `$). We found similar patterns in the multi-species scenario () and the multi-species with opportunity cost scenario (). Although the prioritizations had similar levels of performance, there were different in the their spatial distributions (Figure 3). In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation. Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to generate prioritizations that secure most of the intra-specific genetic variation.

@@

### Limitations/Future directions
We wish to alert practitioners to several limitations associated with our analysis. First, the size of the planning units we used (23 \times 25 km) is much larger than typically used in conservation planning scenarios (approx. 1--5 km^2). We used this resolution because the genetic data was collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because the surrogate variables are spatially auto-correlated. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @rXXX], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time. Overall, the limitations associated with this study are relatively minor.

>>>>

# load .rda file
try(session::restore.session('results/results.rda'))

@@

To understand the relationship between the surrogate-based targets used to generate a prioritizations and the amount of genetic variation secured, we generated a collection of single-species prioritizations. For each species, we generated two sets of prioritizations using increasing environmental and geographic-based targets (respectively). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target). To determine if these surrogate-based prioritizations performed better then random, we generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target). To account for variation in the number of selected planning units in the surrogate-based prioritizations, the number of selected of planning units in the randomly generated prioritizations were generated following their distribution.

@@

For each scenario, we generated three types of prioritizations. We generated prioritizations using only amount-based, both amount- and surrogate-based targets, and both amount- and genetic-based targets. The purely amount-based prioritizations represented those typically generated in conservation planning exercises and serve as our baseline. We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX]. We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the repsective variation was secured for each species. To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations. Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.

@@

## Results
<!-- Surrogacy analysis -->
Prioritizations generated using greater environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1). The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$). Generally, prioritizations needed to capture at least `r NA` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation. Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$). Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation. These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation.

<!-- Scenario analysis -->
The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2). No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$), or under any of the different scenarios (${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$). The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation. Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r NA` $\pm$ `r NA`) and neutral (mean `r NA` $\pm$ `r NA`) genetic variation. These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r NA`$, $P < `r NA`$) or genetic-based targets ($Z = `r NA`$, $P < `r NA`$). We found similar patterns in the multi-species scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$) and the multi-species with opportunity cost scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$). Although the prioritizations had similar levels of performance, there were different in the their spatial distributions (Figure 3). In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation. Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to generate prioritizations that secure most of the intra-specific genetic variation.

@@

### Limitations/Future directions
We wish to alert practitioners to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (approx. 1--5 km^2). We used this resolution because the genetic data was collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because the surrogate variables are spatially auto-correlated. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @rXXX], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time. Overall, the limitations associated with this study are relatively minor.

### 2016-02-21 article/figures.Rmd
+ 4 species for debugging + fixes to article so it compiles + fixes to nmds analyses so they use sparse distances

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
results.SDF <- results.SDF %>% mutate(
		Prioritisation=factor(as.character(Prioritisation), levels=c('Amount','Surrogate','Genetic')),
		Context=factor(as.character(Context), levels=c("single-species (equal costs)", "multi-species (equal costs)", "multi-species (opportunity costs)")),
		Metric=factor(as.character(Metric), c("Adaptive variation", "Neutral variation"))
	)
results.PDF <- results.SDF %>% group_by(
		Prioritisation, Context, Metric
	) %>% mutate(
		letter.height=max(value, na.rm=TRUE)+0.04,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)
	) %>% filter(
		c(TRUE, rep(FALSE, length(Context)-1))
	) %>% data.frame() %>% group_by(Context, Metric) %>% mutate(
		panel.letter=paste0('(',letters[as.integer(Context) + (as.integer(Metric)-1)],')'),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% data.frame() %>% left_join(
		data.frame(
			letter=unname(toupper(cld(full.model.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(full.model.GLHT)$mcletters$Letters)))
		) %>% mutate(Combination=as.character(Combination)),
		by='Combination'
	) %>% data.frame()
results.SDF <- results.SDF %>% mutate(
	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
	Context=factor(
		gsub(' (', '\n(', Context, fixed=TRUE),
		levels=c("Single-species\n(equal costs)",
		"Multi-species\n(equal costs)", 
		"Multi-species\n(opportunity costs)"))
	)
results.PDF <- results.PDF %>% mutate(
	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
	Context=factor(
		gsub(' (', '\n(', Context, fixed=TRUE),
		levels=c("Single-species\n(equal costs)",
		"Multi-species\n(equal costs)", 
		"Multi-species\n(opportunity costs)"))
	)
# make plot
suppressWarnings({
	ggplot(aes(x=Prioritisation,y=value),
		data=results.SDF) +
		geom_boxplot(position=position_dodge(0.9)) +
		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),
			data=results.PDF, position=position_dodge(0.9)) +
		geom_text(aes(x=0.6, y=1.03, label=panel.letter),
			data=results.PDF) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		ylim(c(NA, 1.05)) +
		theme_classic() +
		theme(strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white'),
			panel.border = element_rect(color='black', fill=NA),
			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
		facet_grid(Context ~ Metric)
})
```

>>>>

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogates and adaptive genetic variation. Panel (b) shows that for geographic surrogates and neutral genetic variation. Transparent gray lines denote trends for each species, and black lines show general trends across all species.'}
### prepare data frames for plotting
#env.correlation.PDF <- filter(correlation.DF, Type='Environmental')
#
#%>% left_join(
#		data.frame(
#			letter=unname(toupper(cld(full.model.GLHT)$mcletters$Letters)),
#			Combination=as.character(names(toupper(cld(full.model.GLHT)$mcletters$Letters)))
#		) %>% mutate(Combination=as.character(Combination)),
#		by='Combination'
#	) %>% data.frame()

#geo.correlation.PDF <- filter(correlation.DF, Type='Geographic')
### make plots
#suppressWarnings({
#	p1 <- ggplot() +
#		geom_boxplot(aes(x=Surrogate.target, y=genetic.held, color=Method, group=Method), data=env.correlation.DF) +
#		geom_text(aes(x=Surrogate.target, y=letter.height, label=letter),
#			data=results.PDF, position=position_dodge(0.9)) +
#		geom_text(aes(x=0.6, y=1.03, label='(a)')) +
#		xlab('Environmental variation secured (%)') +
#		ylab('Adaptive genetic\nvariation secured (%)') +
#		theme_classic()
#	p2 <- ggplot() +
#		geom_boxplot(aes(x=Surrogate.target, y=neutral.held, color=Method), data=geo.correlation.DF) +
#		geom_text(aes(x=Surrogate.target, y=letter.height, label=letter),
#			data=results.PDF, position=position_dodge(0.9)) +
#		geom_text(aes(x=0.6, y=1.03, label='(b)')) +
#		xlab('Geographic variation secured (%)') +
#		ylab('Neutral genetic\nvariation secured (%)') +
#		theme_classic()
#	grid.arrange(p1, p2, nrow=1)
# })

plot(1)

```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
results.SDF <- results.SDF %>% mutate(
		Prioritisation=factor(as.character(Prioritisation), levels=c('Amount','Surrogate','Genetic')),
		Context=factor(as.character(Context), levels=c("single-species (equal costs)", "multi-species (equal costs)", "multi-species (opportunity costs)")),
		Metric=factor(as.character(Metric), c("Adaptive variation", "Neutral variation"))
	)
results.PDF <- results.SDF %>% group_by(
		Prioritisation, Context, Metric
	) %>% mutate(
		letter.height=max(value, na.rm=TRUE)+0.04,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)
	) %>% filter(
		c(TRUE, rep(FALSE, length(Context)-1))
	) %>% data.frame() %>% group_by(Context, Metric) %>% mutate(
		panel.letter=paste0('(',letters[as.integer(Context) + (as.integer(Metric)-1)],')'),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% data.frame() %>% left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.model.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.model.GLHT)$mcletters$Letters)))
		) %>% mutate(Combination=as.character(Combination)),
		by='Combination'
	) %>% data.frame()
results.SDF <- results.SDF %>% mutate(
	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
	Context=factor(
		gsub(' (', '\n(', Context, fixed=TRUE),
		levels=c("Single-species\n(equal costs)",
		"Multi-species\n(equal costs)", 
		"Multi-species\n(opportunity costs)"))
	)
results.PDF <- results.PDF %>% mutate(
	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
	Context=factor(
		gsub(' (', '\n(', Context, fixed=TRUE),
		levels=c("Single-species\n(equal costs)",
		"Multi-species\n(equal costs)", 
		"Multi-species\n(opportunity costs)"))
	)
# make plot
suppressWarnings({
	ggplot(aes(x=Prioritisation,y=value),
		data=results.SDF) +
		geom_boxplot(position=position_dodge(0.9)) +
		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),
			data=results.PDF, position=position_dodge(0.9)) +
		geom_text(aes(x=0.6, y=1.03, label=panel.letter),
			data=results.PDF) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		ylim(c(NA, 1.05)) +
		theme_classic() +
		theme(strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white'),
			panel.border = element_rect(color='black', fill=NA),
			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
		facet_grid(Context ~ Metric)
})
```

### 2016-02-21 article/supporting_information.Rmd
+ 4 species for debugging + fixes to article so it compiles + fixes to nmds analyses so they use sparse distances

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=10, fig.cap='Relative support ($BIC$) for the number of populations in each species. Each panel denotes a different species. Each line denotes a different model parametrization. Asterisks indicate the model and number of populations with the most support.'}
# format data
bic.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		curr.MTX <- spp.Mclust.LST[[i]]$mclust$BIC
		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
		attr(curr.MTX, 'dimnames') <- list(
			seq_len(nrow(curr.MTX)),
			attr(curr.MTX, 'modelNames')
		)
		class(curr.MTX) <- 'matrix'
		curr.DF <- curr.MTX %>% as.data.frame %>% 
			mutate(G=attr(curr.MTX, 'G')) %>% 
			gather(Model,BIC,-G)
		curr.DF$BIC=curr.DF$BIC*-1
		return(
			mutate(
				curr.DF,
				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
				best_stat=min(curr.DF$BIC,na.rm=TRUE),
				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
			)
		)
	}
)
# define ggplot2 color wheel function
gg_color_hue <- function(n) {
	hues <- seq(15, 375, length=n+1)
	hcl(h=hues, l=65, c=100)[1:n]
}
# make plot
suppressWarnings({
	ggplot(data=bic.DF) +
		geom_point(aes(y=BIC, x=G, color=Model)) +
		geom_line(aes(y=BIC, x=G, color=Model)) +
		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
		theme_classic() +
		scale_y_reverse()+
		theme(
			strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white', face='italic', size=20)
		) +
		xlab('Number of populations') +
		ylab('BIC') +
		scale_linetype_manual(value=rep(c('solid','dashed'),each=7)) +
		scale_color_manual(value=rep(gg_color_hue(7),2) +
		facet_wrap(~ Species, ncol=4, scales='free_y')
})
```

```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Distribution of populations. Each panel depicts a different species. Squares denote planning units. Points denote the location of sampled individuals, and their colors denote the distribution of populations. Note that points have been slightly jittered so that individuals in close proximity are visible.'}
# prepare data
pop.DF <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(i) {
		curr.pts <- filter(
			spp.sample.PTS@data,
			species==unique(spp.samples.DF$species)[i]
		) %>% mutate(
			species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
			population=spp.Adegenet.LST[[i]]$grp
		)
	}
) %>% do.call(what=rbind)
# make plot
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey75') +
	geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
		group='group'), fill='transparent', color='grey10') +
	geom_point(data=pop.DF, aes_string(x='sample.longitude', y='sample.latitude', 
		color='population'), alpha=0.8, size=3, position='jitter') +
	guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(), 
		axis.title=element_blank(), 
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', face='italic', size=20),
		legend.text= element_text(size=20)
	) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	) +
	facet_wrap(~ species, ncol=4)
```

@@

```{r, echo=FALSE}
knitr::kable(
	format.table(
		filter(
			ldply(
				seq_along(unique(spp.samples.DF$species)), 
				function(i) {
					ldply(
						seq_along(spp.nmds.LST[[i]]),
						function(j) {
						rbind(
							data.frmae(
								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
								Loci='all',
								k=spp.Mclust.LST[[i]]$nmds$ndim,
								Stress=spp.Mclust.LST[[i]]$nmds$stress,
								Converged=spp.Mclust.LST[[i]]$nmds$converged
							),
							data.frame(
								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
								Loci=names(spp.nmds.LST[[i]])[j],
								k=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_integer_, spp.nmds.LST[[i]][[j]]$ndim),
								Stress=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_real_, spp.nmds.LST[[i]][[j]]$stress),
								Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
							)
					})
				}
			),
			!is.na(Stress)
		),
		omit=c('Loci','Converged')
	),
	digits=2,
	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species. The analyses on all loci were used to transform the binary presence/absence polymorphism data to continuous variables to train Gaussian mixture models and ultimately identify adaptive loci. The analyses on adaptive and neutral loci were used to transform the polymorphism data to identify main gradients of this variation.',
	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
	align=c('l', 'c', 'c', 'c', 'c')
)
```

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=10, fig.cap='Relative support ($BIC$) for the number of populations in each species. Each panel denotes a different species. Each line denotes a different model parametrization. Asterisks indicate the model and number of populations with the most support.'}
# format data
bic.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		curr.MTX <- spp.Mclust.LST[[i]]$mclust$BIC
		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
		attr(curr.MTX, 'dimnames') <- list(
			seq_len(nrow(curr.MTX)),
			attr(curr.MTX, 'modelNames')
		)
		class(curr.MTX) <- 'matrix'
		curr.DF <- curr.MTX %>% as.data.frame %>% 
			mutate(G=attr(curr.MTX, 'G')) %>% 
			gather(Model,BIC,-G)
		curr.DF$BIC=curr.DF$BIC*-1
		return(
			mutate(
				curr.DF,
				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
				best_stat=min(curr.DF$BIC,na.rm=TRUE),
				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
			)
		)
	}
)
# define ggplot2 color wheel function
gg_color_hue <- function(n) {
	hues <- seq(15, 375, length=n+1)
	hcl(h=hues, l=65, c=100)[1:n]
}
# make plot
suppressWarnings({
	ggplot(data=bic.DF) +
		geom_line(aes(y=BIC, x=G, color=Model, linetype=Model)) +
		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
		theme_classic() +
		scale_y_reverse()+
		theme(
			strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white', face='italic', size=20)
		) +
		xlab('Number of populations') +
		ylab('BIC') +
		scale_linetype_manual(values=rep(c('solid','dashed'),each=7)) +
		scale_color_manual(values=rep(gg_color_hue(7),2)) +
		facet_wrap(~ Species, ncol=4, scales='free_y')
})
```

```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Distribution of populations. Each panel depicts a different species. Squares denote planning units. Points denote the location of sampled individuals, and their colors denote the distribution of populations. Note that points have been slightly jittered so that individuals in close proximity are visible.'}
# prepare data
pop.DF <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(i) {
		curr.pts <- filter(
			spp.sample.PTS@data,
			species==unique(spp.samples.DF$species)[i]
		) %>% mutate(
			species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
			population=apply(spp.Mclust.LST[[i]]$mclust$z, 1, which.max)
		)
	}
) %>% do.call(what=rbind)
# make plot
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey75') +
	geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
		group='group'), fill='transparent', color='grey10') +
	geom_point(data=pop.DF, aes_string(x='sample.longitude', y='sample.latitude', 
		color='population'), alpha=0.8, size=3, position='jitter') +
	guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(), 
		axis.title=element_blank(), 
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', face='italic', size=20),
		legend.text= element_text(size=20)
	) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	) +
	facet_wrap(~ species, ncol=4)
```

```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Population membership probabilities. Each panel depicts a different species. Each vertical bar represents a different individual, and colors show the probability that it belongs to different populations. This type of plot is also known as a \texttt{Structure} plot.'}
plot(1)
```

@@

```{r, echo=FALSE}
knitr::kable(
	format.table(
		filter(
			ldply(
				seq_along(unique(spp.samples.DF$species)), 
				function(i) {
					ldply(
						seq_along(spp.nmds.LST[[i]]),
						function(j) {
						rbind(
							data.frame(
								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
								Loci='all',
								k=spp.Mclust.LST[[i]]$nmds$ndim,
								Stress=spp.Mclust.LST[[i]]$nmds$stress,
								Converged=spp.Mclust.LST[[i]]$nmds$converged
							),
							data.frame(
								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
								Loci=names(spp.nmds.LST[[i]])[j],
								k=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_integer_, spp.nmds.LST[[i]][[j]]$ndim),
								Stress=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_real_, spp.nmds.LST[[i]][[j]]$stress),
								Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
							)
						)
					})
				}
			),
			!is.na(Stress)
		),
		omit=c('Loci','Converged')
	),
	digits=2,
	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species. The analyses on all loci were used to transform the binary presence/absence polymorphism data to continuous variables to train Gaussian mixture models and ultimately identify adaptive loci. The analyses on adaptive and neutral loci were used to transform the polymorphism data to identify main gradients of this variation.',
	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
	align=c('l', 'c', 'c', 'c', 'c')
)
```
